Dependent Object Types

### 2012-06-30 doc/semester/report.tex
blurb

\documentclass[preprint]{sigplanconf}

@@

\section{Non-Expanding Types and Subtyping Transitivity}

>>>>

\documentclass[preprint,9pt]{sigplanconf}

@@

\subsection{Subtyping Transitivity and Preservation}

Some subtyping transitivity is essential for soundness. Indeed, I show
how to construct a counterexample to preservation from any
counterexample to subtyping transitivity where the three involved
types $S, T, U$ are expressible within a universe $u$ but not
necessarily realizable:

\begin{align*}
\mlnew u \ldots {
\ \mlapp{\abs x \Top x}{
\ \ \mlapp{\abs f {S \tfun U} f}{
\ \ \ \mlapp{\abs f {S \tfun T} f}{
\ \ \ \ \mlapp{\abs f {S \tfun S} f}{
\ \ \  \ \ \abs x S x}}}}}
\end{align*}

The idea is to start with a function from $S \tfun S$ and cast it
successively to $S \tfun T$ then $S \tfun U$. To typecheck the
expression initially, we need to check $S \sub T$ and $T \sub
U$. After some reduction steps, the first few casts vanish, and the
reduced expression casts directly from $S \tfun S$ to $S \tfun U$, so
we need to check $S \sub U$. Thus, we need subtyping transitivity: $S
\sub T$ and $T \sub U$ implies $S \sub U$.

\subsection{Non-Expanding Types and Subtyping Transitivity}

The well-formedness rule $\textsc{tsel-wf$_2$}$ allow a type selection
$p.L$ to refer back to itself in its upper bound. This is useful for
expressing recursive class types and F-bounded abstract types.

Consider the simplest possible type that refers to itself: $T = \Top
\refine z {\Ldecl L \Bot z.L}$. This type is well-formed but
non-expanding.

Non-expanding types are problematic for subtyping transitivity.
